Untwisting the Tetrahelix (v0.5)

### 2017-02-19 doc/TetrahelixUntwist.tex
improving the negative rho situation

\documentclass[11pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage[affil-it]{authblk}
\usepackage[toc,page]{appendix}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{amsthm}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a face-to-face stack of regular tetrahedra.
  Considering the edges of these tetrahedra, the resulting structure is attractive and inherently rigid,
  and therefore interesting to architects, mechanical engineers,
  and robotocists.
  A formula that matches the visually apparent helices forming the outer rails of the tetrahelix is derived which
  is convenient for designers.
  This formula defines the vertices of tetrahelices of varying radius, pitch, and curvature, with the BC helix
  as a special case. 
  An addtional special case is of that of $0$ curvature or rail angle, which generates a \emph{tetrabeam}.
  A particular choice of paramaters defines a novel object, the  minimax member length-difference tetrabeam,
  the \emph{equitetrabeam}.
  Linear interpolation of these parameters with the BC-helix parameters defines a continuum of close-to regular
  tetrahelices of designable curvature, pitch based on a single parameter.
  Utility and use for static and variable geometry truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

The Boerdijk--Coxeter helix\cite{coxeter1985simplicial} (BC helix), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Because architects, structural engineers, and robotocists are inspired
by and follow such mathematical models but can build structures and
machines of differing or even dynamically changing length, it is
useful to develop the mathematics of structure formed from tetrahedra
where we relax regularity.  The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
called the Song-Kwon-Kim\cite{song2003spherical} or turret joint,
which allow many members to come to a single point.  Such machines can
follow purely regual mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

Buckminster Fuller called the BC helix a \emph{tetrahelix}\cite{fuller1982synergetics},
a term now commonly used. In this paper we reserve BC helix to mean the purely regular structre and use \emph{tetrahelix} to refer
to any structure isomophic to a the BC helix, whether regular or not.

\begin{figure}[H] %float with two figures
  \label{series}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up}
\end{figure}

@@

$\theta$ is approximately $131.81$ degrees.
The angle $\theta$ is the rotation of a \emph{each} tetrahedra.
That is, a yellow tetrahedron is rotated slightly more than a $1/3$ of a revolution to match the face of the red tetrahedra.
$3 \theta - 2\pi$ is the apparent rotation of $V_3$ relative to $V_0$.

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession.

This formula defines a helix, but it is not any of the helices of the
BC helix, but rather one that winds three times as rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the by the outer edges or rails of
the BC helix.

It is convenient to have a formula that gives us the nodes of just
each colored helix.
\[
H_{BCcolored}(n,c) = V(3n +c)
\]
where $c \in \{0,1,2\}$ specifies which of the rails is being computed.

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.6392 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra are not lined up perfectly.
It is a famous and interesting result that the pitch is irrational, a BC helix never has two tetrahedra
at precisely the same orientation around the $z$-axis. However, this is inconvenient to designers, who
might prefer a rational pitch. For example, a slight irregularity that led to a pitch of precisely 10 tetrahedra
in one revolution would allow an architect to design a column having a basis and a capital in the same relation to the tetrahedra
they touch.

@@

   In principle in any tetrahelix where the three helices have any
   displacement along the $z$ axis there are 9 distinct edge classes:
   3 rail edges, a one-hop length between eachof 3 rails, and two hop
   length between each of three rails, where the two-hop length is a
   maximum length between rails (which could equal the one-hop
   length.) However, we have already shown the pitches and the rail
   lengths are equal in any optimal tetrahelix.

@@

$H_{general}$ generalizes $H_{continuum}$, but forces the user to select an $d_{\rho}$
which has a sensible radius, so it may be less convenient.

Note that when $\rho = 0$ then $h_{\rho} = 1$, but $r_{\rho}$ is not determined.

@@

Note: $ d_{opt} $ is computed by $\rho, r_{opt}$ via the rail angle equation \eqref{railangle}.

@@

>>>>

\documentclass[11pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage[affil-it]{authblk}
\usepackage[toc,page]{appendix}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a face-to-face stack of regular tetrahedra.
  Considering the edges of these tetrahedra, the resulting structure is attractive and inherently rigid,
  and therefore interesting to architects, mechanical engineers,
  and robotocists.
  A formula that matches the visually apparent helices forming the outer rails of the tetrahelix is derived which
  is convenient for designers.
  This formula defines the vertices of tetrahelices of varying radius, pitch, and curvature, with the BC helix
  as a special case. 
  An addtional special case is of that of $0$ curvature or rail angle, which generates a \emph{tetrabeam}.
  A particular choice of paramaters defines a novel object, the  minimax member length-difference tetrabeam,
  the \emph{equitetrabeam}.
  A formula for computing the radius and z-axis travel that gives an optimal minimax tetrahelix for any rail angle
  between $0$ and $\rho_{bc}$ is given, thus defining a continuum of tetrahelices containing the equitetrabeam and
  BC helix.
  Utility and use for static and variable geometry truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

The Boerdijk--Coxeter helix\cite{coxeter1985simplicial} (BC helix), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Because architects, structural engineers, and robotocists are inspired
by and follow such mathematical models but can build structures and
machines of differing or even dynamically changing length, it is
useful to develop the mathematics of structure formed from tetrahedra
where we relax regularity.  The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
called the Song-Kwon-Kim\cite{song2003spherical} or turret joint,
which allow many members to come to a single point.  Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

Buckminster Fuller called the BC helix a \emph{tetrahelix}\cite{fuller1982synergetics},
a term now commonly used. In this paper we reserve \emph{BC helix} to mean the purely regular structre and use \emph{tetrahelix} to refer
to any structure isomophic to a the BC helix, whether regular or not.

\begin{figure}[H] %float with two figures
  \label{closeup}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up}
\end{figure}

@@

$\theta$ is approximately $131.81$ degrees.
The angle $\theta$ is the rotation of a \emph{each} tetrahedra, not the tetraheda along a rail.  In Figure \ref{closeup},
each tetrahedra has either a yellow, blue, or red outer edge or rail.
That is, a blue-rail tetrahedron is rotated slightly more than a $1/3$ of a revolution to match the face of the yellow tetrahedra.

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

The apparent rotation of tetrhedra of the same outer-edge color ($V_3$ relative from $V_0$ for example,
in \eqref{graycoxer}, is $3 \theta - 2\pi$.

This formula defines a helix, but it is not any of the apparent helices, or rail helices, of the
BC helix, but rather one that winds three times as rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the by the outer edges or rails of
the BC helix.

It is convenient to have a formula that gives us the nodes of just
one rail helix, denoted by $c$:
\[
(\forall n \in \mathbb{Z}, \forall c \in \{0,1,2\} : H_{BCcolored}(n,c) = V(3n +c))
\]

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra are not lined up perfectly.
It is a famous and interesting result that the pitch is irrational, a BC helix never has two tetrahedra
at precisely the same orientation around the $z$-axis. However, this is inconvenient to designers, who
might prefer a rational pitch. For example, a slight irregularity that led to a pitch of precisely 10 tetrahedra
in one revolution would allow an architect to design a column having a basis and a capital in the same relation to the tetrahedra
they touch.

@@

   In principle in any tetrahelix where the three helices have any
   displacement along the $z$ axis there are 9 distinct edge classes:
   3 rail edges, a one-hop length between each of 3 rails, and two hop
   length between each of three rails, where the two-hop length is a
   maximum length between rails (which could equal the one-hop
   length.) However, we have already shown the pitches and the rail
   lengths are equal in any optimal tetrahelix.

@@

$H_{general}$ forces the user to select an $d_{\rho}$
which has a sensible radius.

Note that when $\rho = 0$ then $h_{\rho} = 1$, but $r_{\rho}$ is not determined by
Equation \ref{railangle}.

@@

Note: $r_{opt} $ is not an even function. I think it is only valid for non-negative values of $\rho$, an issue
we must face when we build our continuum.
Note: $ d_{opt} $ is computed by $\rho, r_{opt}$ via the rail angle equation \eqref{railangle}.

@@

Engineers has two fundamental mathamematical tools for constructing a tetrahelix.
They may use
